---
title: "HW3_xl2836"
author: "Xinyi Lin"
date: "10/4/2018"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---
```{r}
library(tidyverse)
library(readxl)
```

# Problem 1

## Question 1

Assuming the mean of $X_i$ is $\mu$.

$$\begin{split}
E(s^2) & =E[\frac{1}{n-1}\sum_{i=1}^n(x_i-x)^2]\\
& =E[\frac{1}{n-1}\sum_{i=1}^n(x_i^2-2\bar{x}x_i+\bar{x}^2)] \\
& =\frac{n}{n-1}E[\frac{1}{n}\sum_{i=1}^n(x_i^2-2\bar{x}x_i+\bar{x}^2)] \\
& =\frac{n}{n-1}\{\frac{1}{n}E[\sum_{i=1}^n(x_i^2)]-2E(\bar{x}\frac{1}{n}\sum_{i=1}^nx_i)+E(\bar{x}^2)\} \\
& =\frac{n}{n-1}\{\frac{1}{n}\sum_{i=1}^nE(x_i^2)-2E(\bar{x}^2)+E(\bar{x}^2)\} \\
& =\frac{n}{n-1}\{\frac{1}{n}\sum_{i=1}^nE(x_i^2)-E(\bar{x}^2)\}
\end{split}$$

As $Var(x_i) = E(x_i^2)-(Ex_i)^2$, we can get $E(x_i^2)=Var(x_i)+(Ex_i)^2=\sigma^2+\mu^2$. As $Var(\bar{x}) = E(\bar{x}^2)-(E\bar{x})^2$, we can get $E(\bar{x}^2)=Var(\bar{x})+(E\bar{x})^2=\frac{\sigma^2}{n}+\mu^2$. Then,
$$\begin{split}
E(s^2) & =\frac{n}{n-1}[\frac{1}{n}\times n(\sigma^2+\mu^2)-\frac{\sigma^2}{n}-\mu^2] \\
& =\frac{n}{n-1}\frac{1}{n}\sigma^2 \\
& =\sigma^2
\end{split}$$

## Question 2
$$\begin{split}
\sum_{i=1}^k\sum_{j=1}^{n_i}(y_{ij}-\bar{\bar{y}})^2 & =\sum_{i=1}^k\sum_{j=1}^{n_i}(y_{ij}^2-2y_{ij}\bar{\bar{y}}+\bar{\bar{y}}^2) \\
& =\sum_{i=1}^k\sum_{j=1}^{n_i}(y_{ij}^2+\bar{\bar{y}}^2)-2\bar{\bar{y}}\sum_{i=1}^k\sum_{j=1}^{n_i}y_{ij} \\
& =\sum_{i=1}^k\sum_{j=1}^{n_i}(y_{ij}^2+\bar{\bar{y}}^2)-2\sum_{i=1}^k\sum_{j=1}^{n_i}\bar{\bar{y}}^2 \\
& =\sum_{i=1}^k\sum_{j=1}^{n_i}(y_{ij}^2-\bar{\bar{y}}^2)
\end{split}$$

$$\begin{split}
\sum_{i=1}^k\sum_{j=1}^{n_i}(y_{ij}-\bar{y_l})^2+\sum_{i=1}^k\sum_{j=1}^{n_i}(\bar{y_l}-\bar{\bar{y}})^2 & =\sum_{i=1}^k\sum_{j=1}^{n_i}[y_{ij}^2-2\bar{y_l}y_{ij}+(\bar{y_l})^2+(\bar{y_l})^2-2\bar{y_l}\bar{\bar{y}}+\bar{\bar{y}}^2] \\
& =\sum_{i=1}^k\sum_{j=1}^{n_i}(y_{ij}^2+\bar{\bar{y}}^2)+\sum_{i=1}^k\sum_{j=1}^{n_i}[2(\bar{y_l})^2-2\bar{y_l}\bar{\bar{y}}+\bar{\bar{y}}^2] \\
& =\sum_{i=1}^k\sum_{j=1}^{n_i}(y_{ij}^2+\bar{\bar{y}}^2)+2\sum_{i=1}^k\sum_{j=1}^{n_i}\bar{y_l}^2-2\sum_{i=1}^k\sum_{j=1}^{n_i}\bar{y_l}y_{ij}-2\sum_{i=1}^k\sum_{j=1}^{n_i}\bar{y_l}\bar{\bar{y}} \\
& =\sum_{i=1}^k\sum_{j=1}^{n_i}(y_{ij}^2+\bar{\bar{y}}^2)+2\sum_{i=1}^k\sum_{j=1}^{n_i}\bar{y_l}^2-2\sum_{i=1}^k(\bar{y_l}\sum_{j=1}^{n_i}y_{ij})-2\bar{\bar{y}}\sum_{i=1}^k(\bar{y_l}\sum_{j=1}^{n_i}1) \\
& =\sum_{i=1}^k\sum_{j=1}^{n_i}(y_{ij}^2+\bar{\bar{y}}^2)+2\sum_{i=1}^k\sum_{j=1}^{n_i}\bar{y_l}^2-2\sum_{i=1}^k\sum_{j=1}^{n_i}\bar{y_l}^2-2\sum_{i=1}^k\sum_{j=1}^{n_i}\bar{\bar{y}}^2 \\
& =\sum_{i=1}^k\sum_{j=1}^{n_i}(y_{ij}^2-\bar{\bar{y}}^2) \\
& =\sum_{i=1}^k\sum_{j=1}^{n_i}(y_{ij}-\bar{\bar{y}})^2
\end{split}$$

# Problem 2

First, we need to import data "HeavySmoke.csv" and "NeverSmoke.csv".

```{r}
heavysmoke_df = read_csv(file = "./data/HeavySmoke.csv")
neversmoke_df = read_csv(file = "./data/NeverSmoke.csv")
```

## Question 1

In order to test wether BMI has changed 6 years after quitting smoking, we need to test the means of `BMI_base` and `BMI_6yrs` are different or not. As we don't know the variance of two samples, we use t test.

Assuming the mean of `BMI_base` is $\mu_b$ and the mean of `BMI_6yrs` is $\mu_6$, the difference between `BMI_base` and `BMI_6yrs` is $d_h$. The samples sizes $n$ is 10.

The null hypothesis $H_0$ : $\mu_b = \mu_6$, the alternative hypothesis $H_1$ : $\mu_b \neq \mu_6$.

```{r}
d_h = heavysmoke_df$BMI_6yrs - heavysmoke_df$BMI_base # get the difference
```

The test statustic :
$$t_{s1}= \frac{\bar{d_h}-0}{s_d/\sqrt{n}} $$ 

```{r}
t_s1 = mean(d_h)*sqrt(10)/sd(d_h)
t_std1 = qt(0.975, 9)
```


$t_{s1}$ is `r round(t_s1, 3)` and $t_{9,0.975}$ is  `r round(t_std1, 3)`. As $t_{s1}$ larger than $t_{9,0.975}$, we reject $H_0$.

Interpreatation: At $\alpha = 0.05$ significant level, we reject $H_0$ and conclude that there is enough evidence to prove that the mean of `BMI_base` $\mu_b$ is different from the mean of `BMI_6yrs` $\mu_6$ and their BMI has changed 6 years after quitting smoking.

## Question 2

Assuming the difference between `BMI_base` and `BMI_6yrs` is never-smoke-group is $d_n$ and difference between $d_h$ and $d_n$ is $d$. The samples sizes $n$ is 10. In order to compare the BMI changes between women that quit smoking and women who never smoked, we use two-sample independent t-test to compare the changes in two groups. First, we need to test the two samples have same variance or not. 

Assuming the variances of two samples are $\sigma_n$ and $\sigma_h$ and the null hypothesis $H_0$ : $\sigma_h = \sigma_s$, the alternative hypothesis $H_1$ : $\sigma_h \neq \sigma_n$.

The test statistic:
$$ F=\frac{{S_1}^2}{{S_2}^2} $$

```{r}
d_n = neversmoke_df$BMI_6yrs - neversmoke_df$BMI_base # get the difference
s_n = sd(d_n)
s_h = sd(d_h)
f = s_n^2/s_h^2
```

As $F$ is `r round(f,3)` is smaller than $F_{9,9,0.975}$ `r round(qf(0.975, 9, 9), 3)` and is larger than $F_{9,9,0.025}$ `r round(qf(0.025, 9, 9), 3)`, we fail to reject $H_0$ at $\alpha = 0.05$ significant level and conclude that there is no significant difference between two variances, so we used two-sample independent t-test with equal variances to compare two differences.

The null hypothesis $H_0$ : $d_h = d_s$, the alternative hypothesis $H_1$ : $d_h \neq d_n$.

The test statistic:
$$ t = \frac{\bar{x_1}-\bar{x_2}}{s\sqrt{(\frac{1}{n_1}+\frac{1}{n_2})}} $$
where $s$ is given by
$$ s^2=\frac{(n_1-1)s_1^2+(n_2-1)s_2^2}{n_1+n_2-2} $$

```{r}
s = sqrt(((10-1)*s_n^2 + (10-1)*s_h^2)/(10+10-2))
t_s2 = (mean(d_n)-mean(d_h))/(s*sqrt(1/10+1/10))
t_std2 = qt(0.975, 18)
```

$t_{s2}$ is `r round(t_s2, 3)` and $t_{9,0.975}$ is `r round(t_std2, 3)`. As absolute value of $t_{s2}$ smaller than $t_{9,0.975}$, we fail to reject $H_0$.

Interpreatation: At $\alpha = 0.05$ significant level, we fail to reject $H_0$ and conclude that there is no enough evidence to prove that the BMI changes between women that quit smoking and women who never smoked are different.

## Question 3

The corresponding 95% CI associated with the difference between changes of two groups $d$ is given by:
$$ \bar{x_1}-\bar{x_2}-t_{n_1+n_2-2,1-\alpha/2}s\sqrt{\frac{1}{n_1}+\frac{1}{n_2}} \leq \bar{x_1}-\bar{x_2} \leq \bar{x_1}-\bar{x_2}+t_{n_1+n_2-2,1-\alpha/2}s\sqrt{\frac{1}{n_1}+\frac{1}{n_2}} $$

```{r}
CI_left = mean(d_n)-mean(d_h) - qt(0.975, 18)*s*sqrt(1/10+1/10)
CI_right = mean(d_n)-mean(d_h) + qt(0.975, 18)*s*sqrt(1/10+1/10)
```

As $n$ is 10 and \alpha is 0.05, the corresponding 95% CI associated with $d$ is [`r round(CI_left, 3)`, `r round(CI_right, 3)`].

## Question 4

Study design:

We can conduct a cohort study. First, we colloect the BMI of people who start to quit smoke and the BMI of them 6 years after they quited smoke. Then we select a group of 100 people from those who able to quit smoke for at least 6 years including 50 women and 50 men that age 50-64. After that, we select a group of 100 people including 50 women and 50 men that had never somke and age 50-64 in the same place as the former group and record their BMI of first and sixth years. At last, we use two sample independent t test to test whether their is difference between the changes of these two groups.

Sample size calculating:

We use the following formula to calculate sample sizes.
$$ n=\frac{(\sigma_1^2+\sigma_2^2)(z_{1-\alpha/2}+z_{1-\beta})^2}{\Delta^2} $$
where $\Delta=|\mu_2-\mu_1|$.

And we can know that $\mu_1=3.0$, $\sigma_1=2.0$, $\mu_2=1.7$, $\sigma_2=1.5$.

```{r}
mu_1 = 3.0
mu_2 = 1.7
sig_1 = 2.0
sig_2 = 1.5
pow_1 = 0.8
pow_2 = 0.9
signiflevel_1 = 0.025
signiflevel_2 = 0.05

sample_size = function(pow, signiflevel){
  numerator = (sig_1^2+sig_2^2)*(qnorm(1-signiflevel/2)+qnorm(pow))^2
  denominator = (mu_1-mu_2)^2
  n = numerator/denominator
  return(n)
}
```

The table of sample sizes are shown below:

  Choice   |        2.5% significance level        |        5% significance level
---------- | ------------------------------------- | ------------------------------------ |
80% power  | `r sample_size(pow_1, signiflevel_1)` | `r sample_size(pow_1, signiflevel_2)`|
90% power  | `r sample_size(pow_2, signiflevel_1)` | `r sample_size(pow_1, signiflevel_2)`|

# Problem 3

First, we need to import data "Knee.csv".

```{r}
Knee_df = read_csv(file = "./data/Knee.csv") %>% 
  janitor::clean_names()
```

## Question 1

The descriptive statistics for Below group is:

```{r}
summary(Knee_df$Below)
```

The descriptive statistics for Average group is:

```{r}
summary(Knee_df$Average)
```

The descriptive statistics for Above group is:

```{r}
summary(Knee_df$Above)
```

## Question 2

```{r}
# tidy data
Knee_aov_df =          
  Knee_df %>% 
  gather(key = "group", value = "time", below:above) %>% 
  filter(!is.na(time))

# get ANOVA table
res_knee = lm(time ~ factor(group), data = Knee_aov_df)   
anova(res_knee)
```

Assuming the mean of Below group is $\mu_b$, the mean of Average group is $\mu_{av}$ and the mean of Above group is $\mu_{ab}$.

The null hypothesis $H_0$ : $\mu_b = \mu_{av} = \mu_{ab}$, the alternative hypothesis $H_1$ : at least two means are not equal.

According to the ANOVA table $F_s$ is 19.28 and $F_{2,22,0.99}$ is `r round(qf(0.99, 2, 22), 2)`. As $F_s$ larger than $F_{2,22,0.99}$, we reject $H_0$.

Interpreatation: At $\alpha = 0.01$ significant level, we reject $H_0$ and conclude that at least two means of Below, Average, Above groups' time are not equal.

## Question 3

### Bonferroni

```{r}
pairwise.t.test(Knee_aov_df$time, Knee_aov_df$group, p.adj = 'bonferroni')
k = 2
```

Bonferroni adjustment:
$$\alpha^* = \frac{\alpha}{\begin{pmatrix} k \\ 2 \\ \end{pmatrix}} $$
In these case, the $\alpha^*$ is `r 0.01*2/(3*2)`, $t_{k-1,1-{\alpha^*}}$ is `r round(qt(1-0.0033, 2), 3)`. As all $t$ in t-test table are smaller than $t_{k-1,1-{\alpha^*}}$, there is no significance different between each group.

### Tukey

```{r}
aov(time ~ factor(group), data = Knee_aov_df) %>% 
  TukeyHSD()
```

### Dunnett

```{r}
below_g = Knee_df$below
average_g = Knee_df$average
above_g = Knee_df$above

DescTools::DunnettTest(list(above_g, average_g, below_g))
```

Both of these three pairwise comparisons can test whether there is significant difference between each two groups in multiple groups. However, we can find out that the results of each test are different. The Bonferroni adjustment can test the difference between every two groups in multiple groups and the result shows that there is no significant difference in every two groups as the Bonferroni adjustment is the strictest. The Tukey adjustment can also test the difference between every two groups in multiple groups and the results shows that only below and average groups have no significant difference. The Dunnett adjustment can test the difference between reference groups(below and average groups) and the control group(above group) and the result shows that there are significant difference between both reference groups and control group.

## Question 4

At $\alpha = 0.05$ significant level, we can conclude that when seperating patients to below, average and above groups, according to Tukey and Dunnett adjustments both below and average groups are significant different from above groups in the time required in physical therapy until successful rehabilitation, which means the time required in physical therapy is associated with physical status once patients' physical therapy is above average.

# Problem 4

## Question 1

```{r}
UCBA_df = as.tibble(datasets::UCBAdmissions) %>%     # import data
  janitor::clean_names()

admit_male =                          # get the number of admitted male in each department
  UCBA_df %>% 
  filter(admit == "Admitted", gender == "Male")

admit_female =                        # get the number of admitted female in each department
  UCBA_df %>% 
  filter(admit == "Admitted", gender == "Female")

reject_male =                          # get the number of rejected male in each department
  UCBA_df %>% 
  filter(admit == "Rejected", gender == "Male")

reject_female =                        # get the number of rejected female in each department
  UCBA_df %>% 
  filter(admit == "Rejected", gender == "Female")

x_m = sum(admit_male$n)
x_f = sum(admit_female$n)

n_f = sum(admit_female$n) + sum(reject_female$n)
n_m = sum(admit_male$n) + sum(reject_male$n)

p_m = x_m/n_m
p_f = x_f/n_f
```

Using the point estimation of the proportions of female and male admitted at Berkeley. The point estimation of the proportions of female $\hat{p_f}$ is `r round(p_f, 3)`, the point estimation of the proportions of male $\hat{p_m}$ is `r round(p_m, 3)`.

Using the following formula to get the 95% confidence interval for the proportions of female and male:
$$ \left(\hat{p}-z_{0.975}\sqrt{\frac{\hat{p}(1-\hat{p})}{n}},\hat{p}+z_{0.975}\sqrt{\frac{\hat{p}(1-\hat{p})}{n}}\right) $$
```{r}
# proportion of female
left_CI_female = p_f - qnorm(0.975)*sqrt(p_f*(1-p_f)/n_f)
right_CI_female = p_f + qnorm(0.975)*sqrt(p_f*(1-p_f)/n_f)

# proportion of male
left_CI_male = p_m - qnorm(0.975)*sqrt(p_m*(1-p_m)/n_m)
right_CI_male = p_m + qnorm(0.975)*sqrt(p_m*(1-p_m)/n_m)
```

By using the above formula, we can get the 95% confidence interval for the proportions of female is ( `r round(left_CI_female,3)`, `r round(right_CI_female,3)` ) and the 95% confidence interval for the proportions of male is ( `r round(left_CI_male,3)`, `r round(right_CI_male,3)` )

According to the mean of two proportions, we can find that The point estimation of the proportions of female $\hat{p_f}$ is lightly smaller than the point estimation of the proportions of male $\hat{p_m}$ as well as the confidence interval which might indicates that the true proportions of female admitted in Berkeley is smaller than the true proportions of male admitted in Berkeley.

## Question 2

The null hypothesis $H_0$ : $p_f = p_m$, the alternative hypothesis $H_1$ : $p_f \neq p_m$. The test statustuc with continuity correction is given by:
$$ z = \frac{|\hat{p_1}-\hat{p_2}|-(\frac{1}{2n_1}+\frac{1}{2n_2})}{\sqrt{\hat{p}\hat{q}(\frac{1}{n_1}+\frac{1}{n_2})}} $$
when $\hat{p} = \frac{n_1\hat{p_1}+n_2\hat{p_2}}{n_1+n_2}$.

We create a function `z.prop` to calculate test statistic. The function is shown below.

```{r}
z.prop = function(x1,x2,n1,n2){
  numerator = abs((x1/n1) - (x2/n2)) - (1/(2*n1)+1/(2*n2))
  p.common = (x1+x2) / (n1+n2)
  denominator = sqrt(p.common * (1-p.common) * (1/n1 + 1/n2))
  z.prop.ris = numerator / denominator
  return(z.prop.ris)
}

z_stat = z.prop(x_f, x_m, n_f, n_m)
```

By calculating, we can know,$z$ is `r round(z_stat, 3)` and $z_{1-\alpha/2}$ is `r round(qnorm(0.975), 3)`. $z$ is larger than $z_{1-\alpha/2}$ and p-value is `r round(pnorm(z_stat), 3)`.

Interpretation: At $\alpha = 0.05$ significant level, we reject $H_0$ and conclude that their are significant difference between the true proportions of female admitted in Berkeley and the true proportions of male admitted in Berkeley.